Efficiently stable presentations from error-correcting codes

### 2025-09-16 daj-revision2/referee-response.txt

I am not fully happy with the revision. Most of my comments and sugges
tions have been taken into account, but not all. Several have been ignored for a
good reason (14 15 20 were completely wrong and 7 was perhaps not relevant),
but it seems that the authors have not got what I meant in my comments 34
and 35.

In 34, I was just asking that you explain in the text of your paper how
Corollary 5.12 improves Theorem 7.14 in [JNV+20a]. What you wrote The
statement is stronger than Theorem 7.14 in [JNV+20a], which involves a game
that has a question with long answers is exactly the kind of thing I was asking.
Would you agree to add it to the text?

A: Ok, we added a sentence right before the statement of Corollary 5.12

@@

In 35, I was just suggesting a different way of phrasing your proof. I guess
that I was implicitely complaining that you were reproducing again the same
argument that you had written many times earlier (and that I called “Lemma 1”
in my first report). And that it would be kind to the reader if you encapsulated
the idea as a lemma so that you can only refer to it each time you use it. Now
that you did it in you new Lemma 2.7, you’d better use it rather than reproving
it. See 21 below.

A: [Michael, you had looked at this. Can you check the referee comment 21 and the lemma they propose (see the pdf file for that) and decide if we want to reformulate as they suggest?]

@@

More importantly, the comments 9 10 13 14 19 21 have not been fully ad-
dressed, or resulted in adding new inaccuracies (see points 2–5, 8, 7, 10, 18). In
particular I fully disagree with your reply to 21, see point 18 below.
Here are some comments. Some are new and some are about previous com-
ments.

1. page 9, Definition 2.4. You made the choice to only discuss what is usually
called flexible stability and not the more restrictive notion where you do
not change de dimension/algebra of the representation. This choice is
understandable because flexible stability is much more common that non-
flexible stability, but maybe it’s worth pointing out the stability results
that are true for the stronger form of non-flexible stability. In particular,
this seems to be the case for Theorem 4.1. This deserves to be written.

A: [Michael, are you able to do this? I can do it but I will have to dig out the results one by one as I am not so familiar]

@@

2. page 9, Claim 2.3. There are squares missing. (3) is certainly correct
(perhaps with a different constant than 3), but I did not follow, in the
proof, how you take into account the renormalization of the trace on N .
(4) is obvious with 5 replaced by 1, because by the trace property, if
X = Ui − w∗Viw,
∥X∥τ = ∥X∗∥τ = ∥U −1
i − w∗V −1
i w∥.

A: Thanks, you are correct. Actually the claim was not needed so we removed it.

### 2025-09-16 daj-revision2/efficient_stability.tex
started some referee edits

Regarding this definition, we state the following elementary claim. \tnote{added this, which seems useful; I'm surprised we didn't need it before?}

\begin{claim}
Suppose that $\{U_i\}$ and $\{V_i\}$ are $(\delta,\mu)$-close. Then the following hold:
\begin{align}
\Es{i\sim \mu} \| wU_i w^* - V_i \|_{\tau^\mN} &\leq 3\delta\;,\label{eq:switch-w}\\
\Es{i\sim \mu} \| U_i^{-1} - w^* V_i^{-1} w \|_{\tau^\mM} &\leq 5\delta\;. \label{eq:switch-inv}
\end{align}
\end{claim}

\begin{proof}
For the first inequality, for any $i$ we have by the triangle inequality 
\begin{align*}
  \| wU_i w^* - V_i \|_{\tau^\mN} &\leq \| w U_i w^* - w w^* V_i w w^* \|_{\tau^\mN} + \|(P-ww^*)V_i ww^* \|_{\tau^\mN} + \|V_i(P-ww^*)\|_{\tau^\mN}\\
  &\leq \| U_i - w^* V_i w\|_{\tau^\mM} + 2 \|(P-ww^*)\|_{\tau^\mN}\;,
\end{align*}
where the second line uses that $w^*$ is a partial isometry. Eq.~\eqref{eq:switch-w} follows. For the second inequality, for any $i$,
\begin{align*}
  \| U_i^{-1} - w^* V_i^{-1} w \|_{\tau^\mM} &=  \| 1_\mM -  U_i w^* V_i^{-1} w \|_{\tau^\mM}\\
  &\leq \|w^* w - w^* w U_i w^* V_i^{-1} w \|_{\tau^\mM} + \|w^*w-1_\mM\|_{\tau^\mM} \\
&\leq \|P - wU_iw^* V_i^{-1}\|_{\tau^\mN}  + \|w^*w-1_\mM\|_{\tau^\mM} \\
&= \|V_i - wU_iw^* \|_{\tau^\mN}  + \|w^*w-1_\mM\|_{\tau^\mM} \;,
\end{align*}
where the first inequality is the triangle inequality and the second uses that $w$ is a partial isometry. 
We conclude by taking the average over $i\sim \mu$ and applying~\eqref{eq:switch-w}.
\end{proof}

@@

  \begin{proof}
    Let $\phi:\mathcal{F}(S)\to \mathcal{U}(\cM)$ be as in the lemma, and let $g$ be a $(\delta,\mu_S)$ close representation of $G$.
   Using the triangle inequality and $g(r)=1_\mN$ and then Lemma~\ref{lem:g-phi}, 
  \begin{align*}
  \Es{r\sim \mu_R}\|\phi(r)-\Id\|_\tau &\leq \Es{r\sim \mu_R}\|\phi(r)-w^*g(r)w\|_\tau + \|w^*w-I_\mM\|_\tau\\
  &\leq \Es{r\sim \mu_R}\sum_{i=1}^{|r|}\| w^* g(r_i)w -\phi(r_i)\|_\tau + \|w^*w-I_\mM\|_\tau +2\delta \\
  &=  \sum_{s\in S} \sum_{r\in R} \mu(R) \sum_{i=1}^{|r|} 1_{s\in \{r_i,r_i^{-1}\}} \| w^* g(r_i)w -\phi(r_i)\|_\tau +   \|w^*w-I_\mM\|_\tau +2\delta\\
  &\leq \Big(\sum_{r\in R} \mu_R(r)|r|^2\Big)\Es{s\sim \mu_S }5 \| w^* g(s)w -\phi(s)\|_\tau +    \|w^*w-I_\mM\|_\tau +2\delta\\
  &\leq 8\delta \Big(\sum_{r\in R} \mu_R(r)|r|^2\Big)\;,
  \end{align*}
  where the factor of $5$ that appeared on the third line accounts for the application of~\eqref{eq:switch-inv} whenever necessary, and for the last line we used $\Es{r\sim \mu_R}|r|^2\geq 1$. 
  \end{proof}

@@

In the next section we detail an important application of this test, Proposition~\ref{prop:dim-test}, which is used in the work~\cite{ji2020mip}. (The present proof somewhat simplifies the one given in~\cite{ji2020mip}, but it keeps the same overall strategy.)

>>>>

Regarding this definition, we state the following elementary claim. \tnote{added this, which seems useful; I'm surprised we didn't need it before?}\tnote{I fixed the claim but actually I don't think it is needed..}

\begin{claim}
Suppose that $\{U_i\}$ and $\{V_i\}$ are $(\delta,\mu)$-close. Then the following hold, where $C$ is a universal constant:
\begin{align}
\Es{i\sim \mu} \| wU_i w^* - V_i \|_{\tau^\mN}^2 &\leq C\delta\;,\label{eq:switch-w}\\
\Es{i\sim \mu} \| U_i^{-1} - w^* V_i^{-1} w \|_{\tau^\mM}^2 &\leq \delta\;. \label{eq:switch-inv}
\end{align}
\end{claim}

\begin{proof}
We first note that by definition
\begin{align*} 
1\geq \tau^\mN(ww^*)  
&= \frac{1}{\tau_\infty(P)}\tau_\infty(ww^*) \\
&= \frac{1}{\tau_\infty(P)}\tau^\mM(w^*w)  \\
&\geq \frac{1}{\tau_\infty(P)}(\tau^\mM(I_\mM)-\delta)\;, 
\end{align*}
from which it follows that $\tau_\infty(P)\geq 1-\delta$.
Now for the first inequality, for any $i$ we have by the triangle inequality 
\begin{align*}
  \| wU_i w^* - V_i \|_{\tau^\mN} &\leq \| w U_i w^* - w w^* V_i w w^* \|_{\tau^\mN} + \|(P-ww^*)V_i ww^* \|_{\tau^\mN} + \|V_i(P-ww^*)\|_{\tau^\mN}\\
  &\leq \frac{1}{\tau_\infty(P)}\| U_i - w^* V_i w\|_{\tau^\mM} + 2 \|(P-ww^*)\|_{\tau^\mN}\\
	&\leq \frac{1}{1-\delta}\| U_i - w^* V_i w\|_{\tau^\mM} + 2\delta \;,
\end{align*}
where the second line uses that $w^*$ is a partial isometry. Eq.~\eqref{eq:switch-w} follows. For the second inequality, for any $i$,
\begin{align*}
  \| U_i^{-1} - w^* V_i^{-1} w \|_{\tau^\mM} &=  \| (U_i^{-1} - w^* V_i^{-1} w)^* \|_{\tau^\mM}\\
	%\| 1_\mM -  U_i w^* V_i^{-1} w \|_{\tau^\mM}\\
  %&\leq \|w^* w - w^* w U_i w^* V_i^{-1} w \|_{\tau^\mM} + \|w^*w-1_\mM\|_{\tau^\mM} \\
%&\leq \|P - wU_iw^* V_i^{-1}\|_{\tau^\mN}  + \|w^*w-1_\mM\|_{\tau^\mM} \\
&= \|U_i - w^* V_i^{-1} w \|_{\tau^\mN} \;, %+ \|w^*w-1_\mM\|_{\tau^\mM} \;,
\end{align*}
%where the first inequality is the triangle inequality and the second uses that $w$ is a partial isometry. 
since $U_i$ and $V_i$ are unitaries. 
%We conclude by taking the average over $i\sim \mu$ and applying~\eqref{eq:switch-w}.
\end{proof}

@@

  \begin{proof}
    Let $\phi:\mathcal{F}(S)\to \mathcal{U}(\cM)$ be as in the lemma, and let $g$ be a $(\delta,\mu_S)$ close representation of $G$.
   Using the triangle inequality and $g(r)=1_\mN$ and then Lemma~\ref{lem:g-phi}, 
  \begin{align*}
  \Es{r\sim \mu_R}\|\phi(r)-\Id\|_\tau &\leq \Es{r\sim \mu_R}\|\phi(r)-w^*g(r)w\|_\tau + \|w^*w-I_\mM\|_\tau\\
  &\leq \Es{r\sim \mu_R}\sum_{i=1}^{|r|}\| w^* g(r_i)w -\phi(r_i)\|_\tau + \|w^*w-I_\mM\|_\tau +2\delta \\
  &=  \sum_{s\in S} \sum_{r\in R} \mu(R) \sum_{i=1}^{|r|} 1_{s\in \{r_i,r_i^{-1}\}} \| w^* g(r_i)w -\phi(r_i)\|_\tau +   \|w^*w-I_\mM\|_\tau +2\delta\\
  &\leq \Big(\sum_{r\in R} \mu_R(r)|r|^2\Big)\Es{s\sim \mu_S } \| w^* g(s)w -\phi(s)\|_\tau +    \|w^*w-I_\mM\|_\tau +2\delta\\
  &\leq 8\delta \Big(\sum_{r\in R} \mu_R(r)|r|^2\Big)\;,
  \end{align*}
  where 
	%the factor of $5$ that appeared on the third line accounts for the application of~\eqref{eq:switch-inv} whenever necessary, and 
	for the last line we used $\Es{r\sim \mu_R}|r|^2\geq 1$. 
  \end{proof}

@@

In the next section we detail an important application of this test, Proposition~\ref{prop:dim-test}, which is used in the work~\cite{ji2020mip}. (The present proof somewhat simplifies the one given in~\cite{ji2020mip}, but it keeps the same overall strategy. \tnote{added for revision 2:}We also note that Corollary~\ref{cor:qld} below is qualitatively stronger than~\cite[Theorem 7.14]{ji2020mip} as the latter applies to a slightly different game which in particular has a pair of questions with long answers, of $O(2^m\log q)$ bits, whereas the tester in Figure~\ref{fig:bRM-tester}, which is analysed in the corollary, has short answers, $O(d)$ bits long. The presence of long answers helps the analysis.)
